{\bf {\Large Four decades of water quality change in the upper San Francisco Estuary}}

### 2016-11-10 sftrends_manu.tex
alpine92 ref added

\begin{singlespace}
\title{{\bf {\Large Four decades of water quality change in the upper San Francisco Estuary}}}
\author{
  {\bf {\normalsize Marcus W. Beck$^1$, David Senn$^2$, Phil Bresnahan$^2$, Emily Novick$^2$, James D. Hagy III$^1$,}}
  \\{\bf {\normalsize Thomas Jabusch$^2$, Phil Trowbridge$^2$}}
  \\\\{\textit {\normalsize $^1$USEPA National Health and Environmental Effects Research Laboratory}}
  \\{\textit {\normalsize Gulf Ecology Division, 1 Sabine Island Drive, Gulf Breeze, FL 32561}}
	\\{\textit {\normalsize Phone: 850-934-2480, Fax: 850-934-2401}}
	\\{\textit {\normalsize Emails: \href{mailto:beck.marcus@epa.gov}{beck.marcus@epa.gov}, \href{mailto:hagy.jim@epa.gov}{hagy.jim@epa.gov}}}
  \\\\{\textit {\normalsize $^2$San Francisco Estuary Institute}}
	\\{\textit {\normalsize 4911 Central Avenue, Richmond, CA 94804}}
	\\{\textit {\normalsize Phone: 510-746-7334, Fax: 510-746-7300}}
	\\{\textit {\normalsize Emails: \href{mailto:davids@sfei.org}{davids@sfei.org}, \href{mailto:philb@sfei.org}{philb@sfei.org}, \href{mailto:emilyn@sfei.org}{emilyn@sfei.org}, \href{mailto:thomasj@sfei.org}{thomasj@sfei.org}}}
  \vspace{1in} 
  \\ Version Date:   Mon Oct 31 12:28:10 2016 -0500
	}
\date{}
\maketitle
\end{singlespace}
\clearpage

@@

Multiple biological and physical factors influence nutrient concentrations at different temporal scales \citep[][, references therein]{Cloern10}.  As a result, relationships between nutrients, time, and flow that could be described by quantitative methods like \ac{wrtds} are espected to be non-linear and complex.  A detailed evaluation of nitrogen dynamics at C10 using \ac{wrtds}, including the effects of flow, is exptected reveal \begin{inparaenum}[1\upshape)]
\item an annual trend independent of a seasonal trend, and 
\item varying flow contribution, either as a difference between predicted and flow-normalized results or changes in the nutrient versus flow relationshps at different annual periods.
\end{inparaenum}
Results from this case study were used to demonstrate the breadth of information that can be obtained from the observed time series with \ac{wrtds}.

@@

Note the decrease in Potamocorbula abundance in 2011, 2012.  These are wet years where abundance/biomass of the clams is driven down by lower salinity.  Contrased wtih the annual chlorophyll trends in the same years, the predicted values are above the flow-normalized trend suggesting an increase in chlorophyll with higher flow.  The potential mechanism is therefore a decrease in clam abundance with high flow that releases phytoplankon from filtration pressure.  This also explains the positive association of chlorophyll with flow in recent years (bottom right dynaplot).

>>>>

\begin{singlespace}
\title{{\bf {\Large Four decades of water quality change in the upper San Francisco Estuary}}}
\author{
  {\bf {\normalsize Marcus W. Beck$^1$, David Senn$^2$, Phil Bresnahan$^2$, Emily Novick$^2$, James D. Hagy III$^1$,}}
  \\{\bf {\normalsize Thomas Jabusch$^2$, Phil Trowbridge$^2$}}
  \\\\{\textit {\normalsize $^1$USEPA National Health and Environmental Effects Research Laboratory}}
  \\{\textit {\normalsize Gulf Ecology Division, 1 Sabine Island Drive, Gulf Breeze, FL 32561}}
	\\{\textit {\normalsize Phone: 850-934-2480, Fax: 850-934-2401}}
	\\{\textit {\normalsize Emails: \href{mailto:beck.marcus@epa.gov}{beck.marcus@epa.gov}, \href{mailto:hagy.jim@epa.gov}{hagy.jim@epa.gov}}}
  \\\\{\textit {\normalsize $^2$San Francisco Estuary Institute}}
	\\{\textit {\normalsize 4911 Central Avenue, Richmond, CA 94804}}
	\\{\textit {\normalsize Phone: 510-746-7334, Fax: 510-746-7300}}
	\\{\textit {\normalsize Emails: \href{mailto:davids@sfei.org}{davids@sfei.org}, \href{mailto:philb@sfei.org}{philb@sfei.org}, \href{mailto:emilyn@sfei.org}{emilyn@sfei.org}, \href{mailto:thomasj@sfei.org}{thomasj@sfei.org}}}
  \vspace{1in} 
  \\ Version Date:   Tue Nov 1 11:08:12 2016 -0500
	}
\date{}
\maketitle
\end{singlespace}
\clearpage

@@

Multiple biological and physical factors influence nutrient concentrations at different temporal scales \citep[][, references therein]{Cloern10}.  As a result, relationships between nutrients, time, and flow that could be described by quantitative methods like \ac{wrtds} are espected to be non-linear and complex.  A detailed evaluation of nitrogen dynamics at C10 using \ac{wrtds}, including the effects of flow, is expected to reveal \begin{inparaenum}[1\upshape)]
\item an annual trend independent of a seasonal trend, and 
\item varying flow contribution to nutrient dynamics, either as a difference between predicted and flow-normalized results or changes in the nutrient versus flow relationship at different annual periods.
\end{inparaenum}
Results from this case study were used to demonstrate the breadth of information that can be obtained from the observed time series with \ac{wrtds}.

@@

Note the decrease in Potamocorbula abundance in 2011, 2012.  These are wet years where abundance/biomass of the clams is driven down by lower salinity.  Contrased wtih the annual chlorophyll trends in the same years, the predicted values are above the flow-normalized trend suggesting an increase in chlorophyll with higher flow.  The potential mechanism is therefore a decrease in clam abundance with high flow that releases phytoplankon from filtration pressure.  This also explains the positive association of chlorophyll with flow in recent years (bottom right dynaplot). See suggestions in \citet{Alpine92} regarding flow/grazer relationships in the Bay.
